Detecting AI-Generated Images using ResNet

### 2022-12-13 paper/paper.tex
Update paper

In order to train a classifier that can distinguish between real and generated images, we need a labeled dataset that consists  of both real and generated images for supervised learning. Datasets of this kind is not common, and we'd like to experiment on the latest state-of-the-art image generation AI models, such as DALL·E 2 and Stable Diffusion. We started with real photos from existing public datasets as our raw datasets. Then, we use Diffusers-based image-to-image model to generate image variations from real photos. This way, we can get a labeled dataset that consists of both real and generated images.

@@

>>>>

In order to train a classifier that can distinguish between real and generated images, we built a labeled dataset that consists of both real and generated images for supervised learning. Datasets of this kind is not common, and we'd like to experiment on the latest state-of-the-art image generation AI models, such as DALL·E 2 and Stable Diffusion. We started with real photos from existing public datasets as our raw datasets. Then, we use Diffusers-based image-to-image model to generate image variations from real photos. This way, we can get a labeled dataset that consists of both real and generated images.

@@

As mentioned, we want to generate variations of the given real-world photos from our raw datasets, and then label both the generated images and the original real-world photos together to build a dataset for training our AI-generated image binary classification model. Our main generative model choice is Stable Diffusion since it's open-source and has a relatively small memory footprint.

Stable Diffusion is originally trained to be a text-to-image model, which means it takes texts as input to describe the images for the model to generate. In order for us to use it as an image-to-image model, we use a pipeline that is similar but swaps the text encoder with the image encoder from CLIP. Since the goal for training CLIP is to maximize the cosine similarities for the image presentation by the image encoder and the text presentation by the text encoder for matching (image, text) pairs, after the CLIP has been trained, the output presentation (vector) from it for matching text and image is very similar, thus the swapped-in image encoder in the Stable Diffusion model we use would play the same role of guiding the diffusion model to generate images similar to the input images. The image encoder from CLIP is used to encode the image into a vector, which is then used as image embedding into the diffusion model to guide it during denoising steps, similar to DALL·E 2's image variation generation API. This way, the model appears to generate variations of the given image that are similar to the them.
